---
title: "est_call_rate_stepwise"
output: 
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# libraries
library(tidyverse)
library(MASS)
library(plyr)
library(broom)

# read in data
df = readRDS("data/processed/proc_acou_photoid.rds")

```

## Upcall{.tabset}
Ok so the 'both' tab is the same as backwards and the 'forward' tab isthe same as the full model. Use tabs labeled 'another approach' or the backwards tab

### full model
```{r}
# upcall
# model
sightf1 = glm.nb(up ~yday+ratio_male_female+foraging_bhv_whale+social_bhv_whale+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))
sightf1
```

### stepwise model 'both'
```{r}
# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(sightf1, direction = "both", 
                       trace = TRUE)

step.modelf1
```

### stepwise model 'backward'
```{r}
# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(sightf1, direction = "backward", 
                       trace = TRUE)

step.modelf1
```

### stepwise model 'forward'
```{r}
# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(sightf1, direction = "forward", 
                       trace = TRUE)

step.modelf1
```

### stepwise model 'forward' another approach
```{r}
# min model
min_sight = glm.nb(up ~
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# max model
max_sight = glm.nb(up ~yday+ratio_male_female+foraging_bhv_whale+social_bhv_whale+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(min_sight, direction = "forward", scope = formula(max_sight),
                       trace = TRUE)

step.modelf1
```

### stepwise model 'both' another approach
```{r}
# min model
min_sight = glm.nb(up ~
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# max model
max_sight = glm.nb(up ~yday+ratio_male_female+foraging_bhv_whale+social_bhv_whale+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(min_sight, direction = "both", scope = formula(max_sight),
                       trace = TRUE)

step.modelf1
```

## Gunshot{.tabset}

### full model
```{r}
# upcall
# model
sightf1 = glm.nb(gs ~yday+ratio_male_female+foraging_bhv_whale+social_bhv_whale+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))
sightf1
```


### stepwise model 'backward'
```{r}
# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(sightf1, direction = "backward", 
                       trace = TRUE)

step.modelf1
```


### stepwise model 'forward' another approach
```{r}
# min model
min_sight = glm.nb(gs ~
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# max model
max_sight = glm.nb(gs ~yday+ratio_male_female+foraging_bhv_whale+social_bhv_whale+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(min_sight, direction = "forward", scope = formula(max_sight),
                       trace = TRUE)

step.modelf1
```

### stepwise model 'both' another approach
```{r}
# min model
min_sight = glm.nb(gs ~
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# max model
max_sight = glm.nb(gs ~yday+ratio_male_female+foraging_bhv_whale+social_bhv_whale+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(min_sight, direction = "both", scope = formula(max_sight),
                       trace = TRUE)

step.modelf1
```

## Mid-feq{.tabset}

### full model
```{r}
# upcall
# model
sightf1 = glm.nb(mf ~yday+foraging_bhv_whale+social_bhv_whale+ratio_male_female+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))
sightf1
```


### stepwise model 'backward'
```{r}
# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(sightf1, direction = "backward", 
                       trace = TRUE)

step.modelf1
```


### stepwise model 'forward' another approach
```{r}
# min model
min_sight = glm.nb(mf ~
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# max model
max_sight = glm.nb(mf ~yday+foraging_bhv_whale+social_bhv_whale+ratio_male_female+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(min_sight, direction = "forward", scope = formula(max_sight),
                       trace = TRUE)

step.modelf1
```

### stepwise model 'both' another approach
```{r}
# min model
min_sight = glm.nb(mf ~
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# max model
max_sight = glm.nb(mf ~yday+foraging_bhv_whale+social_bhv_whale+ratio_male_female+
                  offset(log(rec_duration))
                , data = df, control=glm.control(maxit=100))

# step model of the previous model to see what models should have in table
step.modelf1 <- stepAIC(min_sight, direction = "both", scope = formula(max_sight),
                       trace = TRUE)

step.modelf1
```