---
title: "glm_testing"
output: 
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setup
```{r, include=FALSE}
# libraries
library(ggplot2)
library(tidyverse)
library(lubridate)
library(MASS)
library(mgcv)
library(tree)
library(Hmisc)
library(corrplot)
library(RColorBrewer)
library(DescTools)

# data
df = readRDS("data/processed/proc_acou_photoid.rds")
```

# GLM 
SHOULD USE BOTH STEPWISE, ONLY BACKWARD, ONLY FORWARD, ONLY BASE MODEL(NO INTERACTION TERMS), ETC.

## Estimating number of whales from acoutic data only{.tabset}

### call counts linear
```{r}
# model
sightm = lm(num_sighting ~ up+gs+mf+
               yday+dist+
               up:yday+gs:yday+mf:yday+
               yday:dist+
               dist:up+dist:gs+dist:mf
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call counts linear -2rows
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = lm(num_sighting ~ up+gs+mf+
               yday+dist+
               up:yday+gs:yday+mf:yday+
               yday:dist+
               dist:up+dist:gs+dist:mf
             , data = dfs)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call counts linear -2rows, all int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = lm(num_sighting ~ up*gs*mf*yday*dist
             , data = dfs)

# summary of model
summary(sightm) 

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call counts linear -2rows, no int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = lm(num_sighting ~ up+gs+mf+
               yday+dist+
               yday:dist+
               dist:up+dist:gs+dist:mf
             , data = dfs)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call counts neg. binomial
```{r}
# model
sightm = glm.nb(num_sighting ~ up+gs+mf+
               yday+dist+
               up:yday+gs:yday+mf:yday+
               yday:dist+
               dist:up+dist:gs+dist:mf
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call counts neg. binomial no interaction
```{r}
# model
sightm = glm.nb(num_sighting ~ up+gs+mf+
               yday+dist+
               yday:dist+
               dist:up+dist:gs+dist:mf
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call counts poisson no int, no 2 up days
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(num_sighting ~ up+gs+mf+
               yday+dist+
               yday:dist+
               dist:up+dist:gs+dist:mf
             , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates (call/hour) linear
```{r}
# model
sightm = lm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates (call/hour) linear glm
```{r}
# model
sightm = glm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if base model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates linear -2rows
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = lm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = dfs)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates linear -2rows, all int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = lm(num_sighting ~ up_per_hr*gs_per_hr*mf_per_hr*yday*dist
             , data = dfs)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates linear -2rows, no int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = lm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates (call/hour) neg. binomial
```{r}
# model
sightm = glm.nb(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates (call/hour) neg. binomial -2 rows
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm.nb(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = dfs)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
``` 

### call rates (call/hour) neg. binomial -2 rows, no int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm.nb(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = dfs)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
``` 

### call rates (call/hour) poisson
```{r}
# model
sightm = glm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = df, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates (call/hour) poisson -2 rows
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               up_per_hr:yday+gs_per_hr:yday+mf_per_hr:yday+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### call rates (call/hour) poisson -2 rows, all int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(num_sighting ~ up_per_hr*gs_per_hr*mf_per_hr*yday*dist
             , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```


### call rates (call/hour) poisson -2 rows, no int
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(num_sighting ~ up_per_hr+gs_per_hr+mf_per_hr+
               yday+dist+
               yday:dist+
               dist:up_per_hr+dist:gs_per_hr+dist:mf_per_hr
             , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

## Estimating call rates from visual data only{.tabset}

### upcall call rate, neg binomial
```{r}
# model
sightm = glm.nb(up ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))+
                  yday:dist
                , data = df)

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### upcall call rate, poisson
```{r}
# model
sightm = glm(up ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))+
                  yday:dist
                , data = df, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### upcall call rate, quasipoisson
```{r}
# model
sightm = glm(up ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))+
                  yday:dist
                , data = df, family = 'quasipoisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# # step model
# step.model <- stepAIC(sightm, direction = "both",
#                       trace = FALSE)
# 
# # summary of step model
# summary(step.model)
# 
# # looking to see if model was good or awful
# par(mfrow=c(2,2))
# plot(step.model)
```

### upcall call rate -2rows, poisson
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(up ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))
                , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### upcall call rate, poisson, int
```{r}
# model
sightm = glm(up ~yday+dist+num_sighting+ratio_male_female+
                  social_bhv_whale*foraging_bhv_whale*other_bhv_whale+
                  offset(log(rec_duration))
                , data = df, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### upcall call rate, poisson, int, -2 rows
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(up ~yday+dist+num_sighting+ratio_male_female+
                  social_bhv_whale*foraging_bhv_whale*other_bhv_whale+
                  offset(log(rec_duration))
                , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### upcall call rate, poisson, int, -2 rows, no other bhv
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ]

# model
sightm = glm(up ~yday+dist+num_sighting+ratio_male_female+
                  social_bhv_whale*foraging_bhv_whale+
                  offset(log(rec_duration))
                , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### gunshot call rate, poisson
```{r}
# model
sightm = glm(gs ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))
                , data = df, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### gunshot call rate -2rows, poisson
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ] 

# model
sightm = glm(gs ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))
                , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### mid-freq call rate, poisson
```{r}
# model
sightm = glm(mf ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))
                , data = df, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```

### mid-freq call rate -2rows
```{r}
# remove two high upcall call count rows
dfs <- df[-c(36, 37), ] 

# model
sightm = glm(mf ~ yday+dist+
                  num_sighting+ratio_male_female+
                  social_bhv_whale+foraging_bhv_whale+other_bhv_whale+
                  offset(log(rec_duration))
                , data = dfs, family = 'poisson')

# summary of model
summary(sightm)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(sightm)

PseudoR2(sightm, which = NULL)

# step model
step.model <- stepAIC(sightm, direction = "both",
                      trace = FALSE)

# summary of step model
summary(step.model)

# looking to see if model was good or awful
par(mfrow=c(2,2))
plot(step.model)

PseudoR2(step.model, which = NULL)
```
