---
title: "year_comparison"
output: 
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown for comparing the years for variables of interest. 

```{r, include = FALSE}
# libraries
library(tidyverse)
library(MASS)

# data
df = readRDS("data/processed/proc_acou_photoid.rds")
```

Recall assumptions for anova: normality, equal variance, independent. Kruskal-wallis test (non-parametric anova) is run because some years have very small sample sizes (i.e., 2017) and the data for the most part is not normal.

## Upcall{.tabset}

### count
```{r}
kruskal.test(up~year, data = df)
```

### rate
```{r}
kruskal.test(up_per_hr~year, data = df)
```

### production rate
```{r}
kruskal.test(up_per_hr_per_whale~year, data = df)
```

## Gunshot{.tabset}

### count
```{r}
kruskal.test(gs~year, data = df)
```

### rate
```{r}
kruskal.test(gs_per_hr~year, data = df)
```

### production rate
```{r}
kruskal.test(gs_per_hr_per_whale~year, data = df)
```

## Mid-freq{.tabset}

### count
```{r}
kruskal.test(mf~year, data = df)
```

### rate
```{r}
kruskal.test(mf_per_hr~year, data = df)
```

### production rate
```{r}
kruskal.test(mf_per_hr_per_whale~year, data = df)
```

## Number of whales
```{r}
# anova
sight = aov(num_sighting~year, data = df)
summary(sight)

# non parametric anova
kruskal.test(num_sighting~year, data = df)
```

## Demographics{.tabset}

### adult male
```{r}
kruskal.test(adult_male~year, data = df)
```

### adult female
```{r}
kruskal.test(adult_female~year, data = df)
```

### juvenile male
```{r}
kruskal.test(juvenile_male~year, data = df)
```

### juvenile female
```{r}
kruskal.test(juvenile_female~year, data = df)
```

## Behaviours{.tabset}

### foraging rate
```{r}
kruskal.test(foraging_bhv_whale~year, data = df)
```

### social rate
```{r}
kruskal.test(social_bhv_whale~year, data = df)
```

### other behaviour rate
```{r}
kruskal.test(other_bhv_whale~year, data = df)
```

# Month comparison

## Upcall{.tabset}

### count
```{r}
kruskal.test(up~month, data = df)
```

### rate
```{r}
kruskal.test(up_per_hr~month, data = df)
```

### production rate
```{r}
kruskal.test(up_per_hr_per_whale~month, data = df)
```

## Gunshot{.tabset}

### count
```{r}
kruskal.test(gs~month, data = df)
```

### rate
```{r}
kruskal.test(gs_per_hr~month, data = df)
```

### production rate
```{r}
kruskal.test(gs_per_hr_per_whale~month, data = df)
```

## Mid-freq{.tabset}

### count
```{r}
kruskal.test(mf~month, data = df)
```

### rate
```{r}
kruskal.test(mf_per_hr~month, data = df)
```

### production rate
```{r}
kruskal.test(mf_per_hr_per_whale~month, data = df)
```

## Number of whales
```{r}
# anova
sight = aov(num_sighting~month, data = df)
summary(sight)

# non parametric anova
kruskal.test(num_sighting~month, data = df)
```

## Demographics{.tabset}

### adult male
```{r}
kruskal.test(adult_male~month, data = df)
```

### adult female
```{r}
kruskal.test(adult_female~month, data = df)
```

### juvenile male
```{r}
kruskal.test(juvenile_male~month, data = df)
```

### juvenile female
```{r}
kruskal.test(juvenile_female~month, data = df)
```

## Behaviours{.tabset}

### foraging rate
```{r}
kruskal.test(foraging_bhv_whale~month, data = df)
```

### social rate
```{r}
kruskal.test(social_bhv_whale~month, data = df)
```

### other behaviour rate
```{r}
kruskal.test(other_bhv_whale~month, data = df)
```
