---
title: "beginning_stats"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown for begnning the statistical analysis for sonobuoy honours

Readinging in the data
```{r}
ifile = readRDS("data/processed/proc_acou_photoid.rds")
df1= ifile
```


```{r, include=FALSE}
#Some useful libraries for plotting
library(tidyverse)
library(viridis)
library(ggplot2)
library(dplyr)
library(lubridate)
library(grid)
library(plotly)
library(RColorBrewer)
library(oce)
library(xtable)
library(mapproj)
#library(gganimate)
#source('D:/narw_sonobuoys/functions.R')
```

## Visual stacked plots 

## stacked bargraph using ggplot {.tabset}
### Behaviour
```{r}
# plt = df %>%
#   select(as.numeric("yday"), "year", "foraging", "social", "other_bhv") %>%
#   gather(key = "behaviour" , value = "cval", foraging, social, other_bhv) %>%
#   filter(behaviour %in% c('forgaging', 'social', 'other_bhv')) %>%
#   ggplot(aes(x = id, y = cval, fill = behaviour))+
#   geom_bar(position="stack", stat = 'identity')+
#   scale_fill_viridis(discrete = TRUE, option = "D", 
#                      breaks=c("foraging", "social", "other_bhv"),
#                      labels=c("Feeding", "Socializing","Other Behaviour"))+
#   labs(fill = "Behaviour",
#        x = NULL,
#        y = NULL,
#        title = 'Sighted behaviour within 15 km plus an hour before and after recording')+
#   # facet by year
#   facet_wrap(~year)+
#   ylab('Sighted Behaviour per Deployment')+
#   xlab('')+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), 
#         legend.position="right")+
#   #theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
#   #theme(text = element_text(size=20))+
#   theme(axis.text.y = element_text(colour = "black"))+
#   theme(axis.text.x = element_text(colour = "black"))
# plt

pltb = df1 %>%
  select("yday", "foraging", "social", "other_bhv") %>%
  gather(key = "behaviour" , value = "cval", foraging, social, other_bhv) %>%
  filter(behaviour %in% c('foraging', 'social', 'other_bhv')) %>%
  ggplot(aes(x = as.Date(yday,'1970-01-01'), y = cval, fill = behaviour))+
  geom_bar(position = 'stack', stat = 'identity', width = 2)+
  scale_fill_viridis(discrete = TRUE, option = "E", 
                     breaks=c("foraging", "social", "other_bhv"),
                     labels=c("Feeding", "Socializing","Other Behaviour"))+
  labs(fill = "Behaviour",
       x = NULL,
       y = NULL,
       title = 'Sighted behaviour within 15 km plus an hour before and after recording')+
  # facet by year
  #facet_wrap(~year)+
  ylab('Sighted Whale Behaviour')+
  xlab('Week')+
  scale_x_date(date_breaks="weeks", date_labels="%b-%d")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="bottom")+
  theme(axis.text.x = element_text(), 
        legend.position="right")+
  #theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(text = element_text(size=20))+
  theme(axis.text.y = element_text(colour = "black"))+
  theme(axis.text.x = element_text(colour = "black"))
pltb
```

### Calls
```{r}
# plt = df %>%
#   select(as.numeric("yday"), "year", "up_dur_hr", "mf_dur_hr", "gs_dur_hr") %>%
#   gather(key = "call_type" , value = "cval", up_dur_hr, mf_dur_hr, gs_dur_hr) %>%
#   filter(call_type %in% c('up_dur_hr', 'mf_dur_hr', 'gs_dur_hr')) %>%
#   ggplot(aes(x = id, y = cval, fill = call_type))+
#   geom_bar(position="stack", stat = 'identity')+
#   scale_fill_viridis(discrete = TRUE, option = "D", 
#                      breaks=c("up_dur_hr", "mf_dur_hr", "gs_dur_hr"),
#                      labels=c("Up Call", "Mid-frequency","Gunshot"))+
#   labs(fill = "Call Type",
#        x = NULL,
#        y = NULL,
#        title = 'Calls detected per duration of deployment')+
#   # facet by year
#   #facet_wrap(~year)+
#   ylab('Calls per deployment per duration')+
#   xlab('')+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), 
#         legend.position="right")+
#   #theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
#   #theme(text = element_text(size=20))+
#   theme(axis.text.y = element_text(colour = "black"))+
#   theme(axis.text.x = element_text(colour = "black"))
# plt


pltc = df1 %>%
  select("yday", "year", "up_dur_hr", "mf_dur_hr", "gs_dur_hr") %>%
  gather(key = "call_type" , value = "cval", up_dur_hr, mf_dur_hr, gs_dur_hr) %>%
  filter(call_type %in% c('up_dur_hr', 'mf_dur_hr', 'gs_dur_hr')) %>%
  ggplot(aes(x = as.Date(yday,'1970-01-01'), y = cval, fill = call_type))+
  geom_bar(position="stack", stat = 'identity', width = 2)+
  scale_fill_grey(start = 0, end = .9,  
                     breaks=c("up_dur_hr", "mf_dur_hr", "gs_dur_hr"),
                     labels=c("Upcall", "Mid-frequency","Gunshot"))+
  theme_bw()+
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(fill = "Call Type",
       x = NULL,
       y = NULL,
       title = 'Calls detected per duration of deployment')+
  # facet by year
  #facet_wrap(~year)+
  ylab('Call Rate (call/hour)')+
  xlab('Week')+
  scale_x_date(date_breaks="weeks", date_labels="%b-%d")+
  scale_y_continuous(breaks = seq(0,275,50))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="bottom")+
  theme(axis.text.x = element_text(), 
        legend.position="right")+
  #theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(text = element_text(size=20))+
  theme(axis.text.y = element_text(colour = "black"))+
  theme(axis.text.x = element_text(colour = "black"))
pltc
```

### Demographics
```{r}
# plt = df %>%
#   select(as.numeric("yday"), "year", "juvenile_male", "juvenile_female", "adult_male", "adult_female") %>%
#   gather(key = "age_sex" , value = "cval", juvenile_male, juvenile_female, adult_male, adult_female) %>%
#   #add to make sure character or factor to fix legend and graph colour?
#   filter(age_sex %in% c("juvenile_male", "juvenile_female", "adult_male", "adult_female")) %>%
#   ggplot(aes(x = id, y = cval, fill = age_sex))+
#   geom_bar(position="stack", stat = 'identity')+
#   scale_fill_viridis(discrete = TRUE, option = "D", 
#                      breaks=c("juvenile_male", "juvenile_female", "adult_male", "adult_female"),
#                      labels=c("Juvenile, Male", "Juvenile, Female","Adult, Male", "Adult, Female"))+
#   labs(fill = "Age Class, Sex",
#        x = NULL,
#        y = NULL,
#        title = 'Sightings within 15 km plus an hour before and after recording')+
#   # facet by year
#   #facet_wrap(~year)+
#   ylab('Sightings per deployment duration')+
#   xlab('')+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), 
#         legend.position="right")+
#   #theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
#   #theme(text = element_text(size=20))+
#   theme(axis.text.y = element_text(colour = "black"))+
#   theme(axis.text.x = element_text(colour = "black"))
# plt


pltd = df1 %>%
  select("yday", "year", "juvenile_male", "juvenile_female", "adult_male", "adult_female") %>%
  gather(key = "age_sex" , value = "cval", juvenile_male, juvenile_female, adult_male, adult_female) %>%
  #add to make sure character or factor to fix legend and graph colour?
  filter(age_sex %in% c("juvenile_male", "juvenile_female", "adult_male", "adult_female")) %>%
  ggplot(aes(x = as.Date(yday,'1970-01-01'), y = cval, fill = age_sex))+
  geom_bar(position="stack", stat = 'identity', width = 2)+
  scale_fill_viridis(discrete = TRUE, option = "D", 
                     breaks=c("juvenile_male", "juvenile_female", "adult_male", "adult_female"),
                     labels=c("Juvenile, Male", "Juvenile, Female","Adult, Male", "Adult, Female"))+
  labs(fill = "Age Class, Sex",
       x = NULL,
       y = NULL,
       title = 'Sightings within 15 km plus an hour before and after recording')+
  # facet by year
  #facet_wrap(~year)+
  ylab('Sighted Whales')+
  xlab('Week')+
  theme(axis.text.x = element_text(), 
        legend.position="right")+
  scale_x_date(date_breaks="weeks", date_labels="%b-%d")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="bottom")+
  theme(axis.text.x = element_text(), 
        legend.position="right")+
  #theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(text = element_text(size=20))+
  theme(axis.text.y = element_text(colour = "black"))+
  theme(axis.text.x = element_text(colour = "black"))
pltd
```

## important figures that need to consider!!!!
```{r}

plt = df1 %>%
  ggplot(aes(x=as.Date(yday,'1970-01-01'), y=num_sighting))+
  geom_point()+
  geom_smooth(method = 'lm')
plt

plt = df1 %>%
  ggplot(aes(x=as.Date(yday,'1970-01-01'), y=up_dur_hr))+
  geom_point()
plt


plt = df1 %>%
  ggplot(aes(x=num_sighting, y=up_dur_hr))+
  geom_point()+
  geom_smooth(method = 'lm')
plt

```

## Maps
sighting size over space, colour is day and size is number animals sighted
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-61.75), ylim = c(47,49))+
  
  # land and bathymetry
   geom_polygon(data = nam, aes(x = long, y = lat, group = group),
               fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = num_sighting, 
                            colour = yday, shape =as.character(year)))+
  #, alpha = 0.7)+
    scale_fill_manual(values = c("black",'red'))+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

## Maps with calls and behaviour as size {.tabset}
### Up calls 
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-62.7), ylim = c(47.2,48.5))+
  
  # land and bathymetry
   geom_polygon(data = nam, aes(x = long, y = lat, group = group),
              fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = up_dur_hr, colour = yday))+
    scale_radius(range=c(3,8))+
  
  # facet by year
  facet_wrap(~year)+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

### Gunshots
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-62.7), ylim = c(47.2,48.5))+
  
  # land and bathymetry
  geom_polygon(data = nam, aes(x = long, y = lat, group = group),
               fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = gs_dur_hr, colour = yday))+
  scale_radius(range=c(3,8))+
    
  # facet by year
  facet_wrap(~year)+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

### Mid-frequency
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-62.7), ylim = c(47.2,48.5))+
  
  # land and bathymetry
   geom_polygon(data = nam, aes(x = long, y = lat, group = group),
                fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = mf_dur_hr, colour = yday))+
  scale_radius(range=c(3,8))+
  
  # facet by year
  facet_wrap(~year)+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

### Feeding behaviour
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-62.7), ylim = c(47.2,48.5))+
  
  # land and bathymetry
   geom_polygon(data = nam, aes(x = long, y = lat, group = group),
               fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = foraging, colour = yday))+
  scale_radius(range=c(3,8))+
  
  # facet by year
  facet_wrap(~year)+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

### Social behaviour
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-62.7), ylim = c(47.2,48.5))+
  
  # land and bathymetry
   geom_polygon(data = nam, aes(x = long, y = lat, group = group),
                fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = social, colour = yday))+
  scale_radius(range=c(3,8))+
  
  # facet by year
  facet_wrap(~year)+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

### Other behaviour
```{r}
# read in map data
load('data/processed/map_data.rda')

# city positions
cities = data.frame(
  name = c("Shippagan", "Percé"),
  lat = c(47.7439, 48.5244),
  lon = c(-64.7057, -64.2127)
)

# gsl map
gsl = ggplot() + 
  coord_map(xlim = c(-65,-62.7), ylim = c(47.2,48.5))+
  
  # land and bathymetry
   geom_polygon(data = nam, aes(x = long, y = lat, group = group),
               fill = "darkgrey", color = NA) +
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-50),colour="grey", size=0.3)+
  # geom_contour(data = bf, aes(x=x, y=y, z=z),
  #               breaks=c(-100),colour="grey", size=1)+
  # 
  # city labels
  geom_point(data = cities, aes(x = lon, y = lat), color = "darkslategrey", size = 1)+
  geom_text(data = cities, aes(x = lon, y = lat, label = name), color =
              "darkslategrey", nudge_y = 0.07, size = 3)+
  
  # sonobuoy positions
  geom_point(data = df1, aes(x = lon, y = lat, size = other_bhv, colour = yday))+
  scale_radius(range=c(3,8))+
  
  # facet by year
  facet_wrap(~year)+
  
  # formatting
  ylab("Longitude (N)")+xlab("Latitude (W)")+
  theme_bw()+
  theme(legend.title = element_blank(), 
        legend.position = 'bottom', 
        legend.direction = "horizontal",
        panel.spacing = unit(2, "lines"))+
  theme(text = element_text(size=15))+
  theme(axis.text.x = element_text(colour = "black"))+
  theme(axis.text.y = element_text(colour = "black"))

gsl

```

## Seasonality in behaviour

**Question:** Does feeding decrease while socializing behaviour increases over the course of the season?

``` {r}
# ggplot()+
#   geom_point(data = df, aes(x = yday, y = feed), color = 'black')+
#   geom_point(data = df, aes(x = yday, y = social), color = 'blue')
ifile = readRDS("data/processed/proc_acou_photoid.rds")
df1= ifile
library(tidyverse)
library(viridis)
library(ggplot2)
library(dplyr)
library(lubridate)
library(grid)
library(plotly)
library(RColorBrewer)
library(oce)
library(xtable)
library(mapproj)
plt = df1 %>%
  select("yday", "year", "foraging", "social") %>%
  gather(key = "behaviour" , value = "cval", foraging, social) %>%
  filter(behaviour %in% c('foraging', 'social')) %>%
  ggplot(aes(x = as.Date(yday,'1970-01-01'), y = cval))+
  geom_point(aes(colour = behaviour), size = 3)+
  geom_smooth(aes(color = behaviour, fill = behaviour), method = "lm",
              se = TRUE, fullrange = TRUE, show.legend = FALSE)+
  scale_colour_viridis(discrete = TRUE, option = "D",
                       breaks=c("foraging", "social"),
                       labels=c("Feeding", "Socalizing"))+
  labs(colour = "Behaviour Type",
       x = NULL,
       y = NULL)+
  scale_x_date(date_breaks="weeks", date_labels="%b-%d")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="bottom")+
  theme(axis.text.x = element_text(), 
        legend.position="right")+
  theme(text = element_text(size=20))+
  theme(axis.text.y = element_text(colour = "black"))+
  theme(axis.text.x = element_text(colour = "black"))+
  ylab('Behaviour Occurance')+
  xlab('Week')

plt
```

**Answer:** Yes. Feeding seems to happen early, and socializing happens later. The trend is pretty weak and only convincing when you lump years.

## Seasonality in calling

**Question:** Does calling behaviour increase over the course of the season?

``` {r}
ifile = readRDS("data/processed/proc_acou_photoid.rds")
df1= ifile
library(tidyverse)
library(viridis)
library(ggplot2)
library(dplyr)
library(lubridate)
library(grid)
library(plotly)
library(RColorBrewer)
library(oce)
library(xtable)
library(mapproj)
# ggplot()+
#   geom_point(data = df, aes(x = yday, y = up_dur_hr), color = 'black')+
#   geom_point(data = df, aes(x = yday, y = mf_dur_hr), color = 'blue')+
#   geom_point(data = df, aes(x = yday, y = gs_dur_hr), color = 'red')

# ggplot()+
#   geom_point(data = df, aes(x = yday, y = up_dur_hr), color = 'black')
# 
# ggplot()+
#   geom_point(data = df, aes(x = yday, y = mf_dur_hr), color = 'blue')
# 
# ggplot()+
#   geom_point(data = df, aes(x = yday, y = gs_dur_hr), color = 'red')

plt = df1 %>%
  select("yday", "year", "up_dur_hr", "mf_dur_hr", "gs_dur_hr") %>%
  gather(key = "call_type" , value = "cval", up_dur_hr, mf_dur_hr, gs_dur_hr) %>%
  filter(call_type %in% c('up_dur_hr', 'mf_dur_hr', 'gs_dur_hr')) %>%
  ggplot(aes(x = as.Date(yday,'1970-01-01'), y = cval))+
  geom_point(aes(colour = call_type), size = 3)+
  scale_colour_viridis(discrete = TRUE, option = "D",
                   breaks=c("up_dur_hr", "mf_dur_hr", "gs_dur_hr"),
                   labels=c("Up call", "Mid-frequency","Gunshot"))+
  #facet_wrap(~call_type)+
  labs(colour = "Call Type",
       x = NULL,
       y = NULL)+
  theme(text = element_text(size=20))+
  scale_x_date(date_breaks="weeks", date_labels="%b-%d")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="bottom")+
  theme(axis.text.x = element_text(), 
        legend.position="right")+
  theme(axis.text.y = element_text(colour = "black"))+
  theme(axis.text.x = element_text(colour = "black"))+
  ylab('Call Rate (call/hour)')+
  xlab('Week')
plt

```

**Answer:** Yes. Not with upcalls, but with gunshots and MF

## Sightings vs call counts

**Question:** Does calling rate change over the course of the season?

```{r}
ifile = readRDS("data/processed/proc_acou_photoid.rds")
df1= ifile
library(tidyverse)
library(viridis)
library(ggplot2)
library(dplyr)
library(lubridate)
library(grid)
library(plotly)
library(RColorBrewer)
library(oce)
library(xtable)
library(mapproj)

# df %>%
#   mutate(call_rate = sum_calls_dur*60*60/num_sighting) %>%
#   ggplot()+
#   geom_point(aes(x = yday, y = call_rate), color = 'black')

plt = df1 %>%
  mutate(call_rate = sum_calls_dur*60*60/num_sighting) %>%
  ggplot()+
  geom_point(aes(x = as.Date(yday,'1970-01-01'), y = call_rate), color = 'black', size = 3)+
  geom_smooth(aes(x = as.Date(yday,'1970-01-01'), y = call_rate), method = "lm", color='black',
              se = TRUE, fullrange = TRUE)+
  scale_x_date(date_breaks="weeks", date_labels="%b-%d")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="bottom")+
  theme(text = element_text(size=20))+
  theme(axis.text.y = element_text(colour = "black"))+
  theme(axis.text.x = element_text(colour = "black"))+
  ylab('Call Rate per Sighting (call/hour/whale)')+
  xlab('Week')
plt

```

**Answer:** Yes! Calls per hour per whale appears to increase over the course of the season. Note that this is a sum of all call types.

**Question:** Does upcall calling rate change over the course of the season?

```{r}

df1 %>%
  mutate(call_rate = up_dur_hr/num_sighting) %>%
  ggplot()+
  geom_point(aes(x = yday, y = call_rate), color = 'black')

```

**Answer:** Maybe. There is a bit of a decline, but also several large outliers. Not nearly as clear as when using the other two call types.

## Gunshot vs MF

**Question:** Is there a consistent ratio between gunshots and mf?

```{r}
df1 %>%
  mutate(gm_ratio = num_mf / num_gs) %>%
  ggplot()+
  geom_point(aes(x = yday, y = gm_ratio))
```

**Answer:** Not really. Maybe around 4-5 for some deployments, but not all and highly variable...

**Question:** Is there a consistent ratio between mf and upcalls? Could this be used to detect/predict socializing behaviour?

```{r}
df1 %>%
  mutate(gm_ratio = num_mf / num_up) %>%
  ggplot()+
  geom_point(aes(x = yday, y = gm_ratio, color = social))
```

**Answer:** Yes! Make sure to double check the data - maybe too many points??

## Abundance vs seasonality

**Question:** Does abundance change over the season?

```{r}
df1 %>%
  ggplot()+
  geom_point(aes(x = yday, y = num_sighting))
```

**Answer:** Likely not...

## Demographic vs seasonality

**Question:** Does male:female ratio change over the season?

```{r}
df1 %>%
  ggplot()+
  geom_point(aes(x = yday, y = 1/ratio_female_male))
```

**Answer:** Looks like it's more variable initially and converges over the course of the season




## Pair plots for identifying patterns
Making the pairs plots sighting and calls
```{r}
 dfs = subset(df1, select=c('num_sighting',
                           'up_dur_hr',
                           'gs_dur_hr',
                           'mf_dur_hr',
                           'yday',
                           'juvenile_female',
                           'juvenile_male',
                           'adult_male',
                           'adult_female'))
 pairs(dfs)
```

calls to behaviour
```{r}
dfs = subset(df1, select=c('up_dur_hr',
                          'gs_dur_hr',
                          'mf_dur_hr',
                          'yday',
                          'foraging',
                          'social',
                          'other_bhv',
                          'social_sag',
                          'social_notsag',
                          'entg_eff',
                          'mom_calf'))
pairs(dfs)
```

2017 the difference between dep 19 and 20 from the rest
looking at behaviour in 2017 with calls
```{r}
# library(tidyverse)
# library(dplyr)
# yr17 = df %>%
#   filter(year == '2017')
# yr2017s = subset(yr17, select = c('up_dur_hr',
#                           'gs_dur_hr',
#                           'mf_dur_hr',
#                           'SUB_FD',
#                           'DFCN',
#                           'BOD_CNT',
#                           'SAG',
#                           'ROLL',
#                           'HD_LFT'))
# pairs(yr2017s)
```

looking at behaviour in 2018 with calls
```{r}
# yr18 = df %>%
#   filter(year == '2018')
# yr2018s = subset(yr18, select = c('up_dur_hr',
#                           'gs_dur_hr',
#                           'mf_dur_hr',
#                           'SUB_FD',
#                           'DFCN',
#                           'BOD_CNT',
#                           'SAG',
#                           'ROLL',
#                           'HD_LFT'))
# pairs(yr2018s)
```

## t-test for duration
```{r}
# shapiro.test(df$duration)
# d = sum(df$duration)
# meand = d/18
# meand = meand/60/60
# t.test(df$duration/60/60, conf.int = 0.95)
# boxplot(df$duration/60/60)
```

## POOLED t-Tests for comparing all variables of interest among the two years
are number of whales similar among years?
t-test on means for comparing sighted whales per year
```{r}
# sight17 = df1 %>%
#   filter(year == '2017') %>%
#   select(num_sighting)
#    
# sight18 = df1 %>%
#   filter(year == '2018') %>%
#   select(num_sighting)
#  
# 
# t.test(sight17, sight18, var.equal = TRUE, conf.level = 0.95)
# # b1 = boxplot(sight17)
# # b2 = boxplot(sight18)

```
thus yes, because they are not statisitcally significant 

are the number of gs call similar among years?
t-test on means for comparing gs calls per year
```{r}
# c17 = df1 %>%
#   filter(year == '2017') %>%
#   select(gs_dur_hr)
# 
# c18 = df1 %>%
#   filter(year == '2018') %>%
#   select(gs_dur_hr)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant 

are the number of up call similar among years?
t-test on means for comparing up calls per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(up_dur_hr)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(up_dur_hr)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant 

are the number of mf call similar among years?
t-test on means for comparing mf calls per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(mf_dur_hr)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(mf_dur_hr)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# # b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant 

are the feeding behaviour reports similar among years?
t-test on means for comparing feeding per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(feed)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(feed)
#  
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the social behaviour reports similar among years?
t-test on means for comparing social per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(social)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(social)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the other behaviour reports similar among years?
t-test on means for comparing feeding per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(other_behv)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(other_behv)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# # b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the observed ages similar among years?
t-test on means for comparing ages per year - juv
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(sum_juvenile)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(sum_juvenile)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the observed ages similar among years?
t-test on means for comparing ages per year - adult
```{r}
# df2=df%>%
#   mutate(sum_adult=adult_female+adult_male)
# 
# c17 = df2 %>%
#   filter(year == '2017') %>%
#   select(sum_adult)
# 
# c18 = df2 %>%
#   filter(year == '2018') %>%
#   select(sum_adult)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# # b1 = boxplot(c17)
# b2 = boxplot(c18)
```
significant for 0.05 = alpha, this is because it was significant for the females, see t-test below of just adult females 

are the observed ages similar among sexes?
t-test on means for comparing ages per year - female
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(sum_female)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(sum_female)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# # b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the observed ages similar among years?
t-test on means for comparing ages per year - male
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(sum_male)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(sum_male)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the observed sex-ages similar among years?
t-test on means for comparing sex-ages per year - adult female
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(adult_female)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(adult_female)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
significant if alpha = 0.05


are the observed sex-ages similar among years?
t-test on means for comparing sex-ages per year - adult male
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(adult_male)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(adult_male)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes because they are not statisitcally significant

are the durations similar among years?
t-test on means for comparing durations per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   select(duration)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   select(duration)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus yes, because they are not statisitcally significant

are the call rates similar among years?
t-test on means for comparing call rates per year
```{r}
# c17 = df %>%
#   filter(year == '2017') %>%
#   mutate(call_rate = sum_calls_dur*60*60/num_sighting)%>%
#   select(call_rate)
# 
# c18 = df %>%
#   filter(year == '2018') %>%
#   mutate(call_rate = sum_calls_dur*60*60/num_sighting) %>%
#   select(call_rate)
# 
# t.test(c17, c18, var.equal = TRUE, conf.level = 0.95)
# # 
# # b1 = boxplot(c17)
# b2 = boxplot(c18)
```
thus, yes because they are not statsitcally significant 

## Regressions for seasonality 
basic regressions with season as the x-asis

are sightings similar over the season?
```{r}
# lm = lm(df$num_sighting~df$yday)
# anova(lm) 
# summary(lm)
```

are gs_dur_hr calls similar over the season?
```{r}
# lm = lm(df$gs_dur_hr~df$yday)
# anova(lm) 
# summary(lm)
```
NO, significant

are up_dur_hr calls similar over the season? 
```{r}
# lm = lm(df$up_dur_hr~df$yday)
# anova(lm) 
# summary(lm)
```

are mf_dur_hr calls similar over the season?
```{r}
# lm = lm(df$mf_dur_hr~df$yday)
# anova(lm) 
# summary(lm)
```
NO, significant

are feed behaviour similar over the season?
```{r}
# lm = lm(df$feed~df$yday)
# anova(lm) 
# summary(lm)
```
NO, significant

are social behaviour similar over the season?
```{r}
# lm = lm(df$social~df$yday)
# anova(lm) 
# summary(lm)
```
NO, significant

are other behaviour similar over the season?
```{r}
# lm = lm(df$other_behv~df$yday)
# anova(lm) 
# summary(lm)
```

are age-juv similar over the season?
```{r}
# lm = lm(df$sum_juvenile~df$yday)
# anova(lm) 
# summary(lm)
```

are age-adult similar over the season?
```{r}
# df= df %>%
#   mutate(sum_adult = adult_female+adult_male)
# lm = lm(df$sum_adult~df$yday)
# anova(lm) 
# summary(lm)
```

are sex-female similar over the season?
```{r}
# lm = lm(df$sum_female~df$yday)
# anova(lm) 
# summary(lm)
```

are sex-male similar over the season?
```{r}
# lm = lm(df$sum_male~df$yday)
# anova(lm) 
# summary(lm)
```

are adult female similar over the season?
```{r}
# lm = lm(df$adult_female~df$yday)
# anova(lm) 
# summary(lm)
```

are adult male similar over the season?
```{r}
# lm = lm(df$adult_male~df$yday)
# anova(lm) 
# summary(lm)
```

are durations similar over the season?
```{r}
# lm = lm(df$duration~df$yday)
# anova(lm) 
# summary(lm)
```

are call rates similar over the season?
```{r}
# df = df %>%
#   mutate(call_rate = sum_calls_dur*60*60/num_sighting)
# lm = lm(df$call_rate~df$yday)
# anova(lm)
# summary(lm)
```
NO, significant

## Regressions over space (latitude)
basic regressions with latitude as the x-asis

are sightings similar over latitude?
```{r}
# lm = lm(df$num_sighting~df$lat)
# anova(lm) 
# summary(lm)
```

are gs_dur_hr calls similar over latitude?
```{r}
# lm = lm(df$gs_dur_hr~df$lat)
# anova(lm) 
# summary(lm)
```

are up_dur_hr calls similar over latitude?
```{r}
# lm = lm(df$up_dur_hr~df$lat)
# anova(lm) 
# summary(lm)
```

are mf_dur_hr calls similar over latitude?
```{r}
# lm = lm(df$mf_dur_hr~df$lat)
# anova(lm) 
# summary(lm)
```

are feed behaviour similar over latitude?
```{r}
# lm = lm(df$feed~df$lat)
# anova(lm) 
# summary(lm)
```

are social behaviour similar over latitude?
```{r}
# lm = lm(df$social~df$lat)
# anova(lm) 
# summary(lm)
```

are other behaviour similar over latitude?
```{r}
# lm = lm(df$other_behv~df$lat)
# anova(lm) 
# summary(lm)
```

are age-juv similar over latitude?
```{r}
# lm = lm(df$sum_juvenile~df$lat)
# anova(lm) 
# summary(lm)
```

are age-adult similar over latitude?
```{r}
# df= df %>%
#   mutate(sum_adult = adult_female+adult_male)
# lm = lm(df$sum_adult~df$lat)
# anova(lm) 
# summary(lm)
```

are sex-female similar over latitude?
```{r}
# lm = lm(df$sum_female~df$lat)
# anova(lm) 
# summary(lm)
```

are sex-male similar over latitude?
```{r}
# lm = lm(df$sum_male~df$lat)
# anova(lm) 
# summary(lm)
```

are adult female similar over latitude?
```{r}
# lm = lm(df$adult_female~df$lat)
# anova(lm) 
# summary(lm)
```

are adult male similar over latitude?
```{r}
# lm = lm(df$adult_male~df$lat)
# anova(lm) 
# summary(lm)
```

are durations similar over latitude?
```{r}
# lm = lm(df$duration~df$lat)
# anova(lm) 
# summary(lm)
```

are call rates similar over latitude?
```{r}
# df = df %>%
#   mutate(call_rate = sum_calls_dur*60*60/num_sighting)
# lm = lm(df$call_rate~df$lat)
# anova(lm)
# summary(lm)
```



## Regressions that were significant 
basic regressions and anova tests that were significant 
```{r}
# # i don like this tho, many whales does not imply all each making a single #sound
# lm = lm(df$sum_calls~df$num_sighting)
#  anova(lm) #SIGNIFICANT now, wasnt before added the rest of dep 16 data
#  summary(lm) #Rsquare = 0.2334, 0.1823
#  cor(x=df$num_sighting, y=df$sum_calls)
#  
#  lm = lm(df$sum_calls~df$adult_male)
#  anova(lm) #SIGNIFICANT!!!!!! 
#  summary(lm) #Rsquared = 0.3009, 0.2543
#  
# #i think this is the key to saying that the number of males depends on the
# # the number of mf calls there are 
#  lm = lm(df$num_mf~df$adult_male)
#  anova(lm) #SIGNIFICANT!!!!!
#  summary(lm) #Rsquared = 0.3419, 0.298
#  #plot(lm)
#  lm = lm(df$mf_dur_hr~df$adult_male)
#  anova(lm) #SIGNIFICANT!!!!!
#  
#  lm = lm(df$mf_dur_hr~df$gs_dur_hr)
#  #plot(lm)
#  anova(lm) #SIGNIFICANT!!!!!!
#  summary(lm)
#  
#  lm = lm(df$mf_dur_hr~df$adult_male+df$juvenile_male+df$num_sighting+df$gs_dur_hr)
#  anova(lm) #SIGNIFICANT!!!!!!
#  
#  lm2a = lm(df$num_mf~df$adult_male+df$juvenile_male)
#  anova(lm2a) #BOTH ARE SIGNIFICANT!!!!!
# #males are not making midfrequency calls
#  summary(lm2a) #rsquared = 0.4977, 0.4259 
#  #plot(lm2a)
#  
#  lm2b = lm(df$mf_dur_hr~df$adult_male)
#  anova(lm2b) #SIGNIFICANT!!!!!!
#  summary(lm2b)
#  
#  lm4 = lm(df$mf_dur_hr~df$num_sighting)
#  anova(lm4) #SIGNIFICANT!!!!well duh most of the whales are male
#  summary(lm4)
```
